Revisiting the hydrodynamic modulation of short surface waves by longer waves

### 2024-10-15 paper.tex
Final edits for submission on 10/15/2024

\begin{abstract}
Hydrodynamic modulation of short ocean surface waves by longer ambient waves is
a well-known ocean surface process that affects remote sensing, the
interpretation of \textit{in situ} wave measurements, and numerical wave
forecasting.
In this paper, we revisit the linear (Airy) wave theory and derive higher-order
steady solutions for the change of short-wave wavenumber, action density, and
gravitational acceleration due to the presence of longer waves.
We validate the analytical solutions with numerical simulations of the full wave
crest and action conservation equations.
The nonlinear analytical solutions of short-wave wavenumber, amplitude, and
steepness modulation significantly deviate from the linear analytical solutions
of \citet{longuet1960changes}, and are similar to the nonlinear numerical
solutions by \citet{longuet1987propagation} and \citet{zhang1990evolution}.
The short-wave steepness modulation is attributed to be 5/8 due to the
wavenumber, 1/4 due to the wave action, and 1/8 due to the effective gravity.
We further examine the result of \citet{peureux2021unsteady} who found through
numerical simulations that the short-wave modulation grows unsteadily with
each long-wave passage.
We show that this unsteady growth only occurs only for homogeneous initial
conditions as a special case and does not generally occur, for example in more
realistic long-wave groups.
The combined analytical and numerical framework presented here allows
quantifying the importance of various terms in the full wave conservation
equations.
The proposed steady solutions are a good approximation of the fully
nonlinear numerical solutions in long-wave steepness up to $\approx 0.2$.
As a corollary, except for a subset of initial conditions, the solutions to
the fully nonlinear crest-action conservation equations are mostly steady in
the reference frame of the long waves.
\end{abstract}

@@

Short waves that ride on the surface of the longer waves move in an accelerated
reference frame (due to the centripetal acceleration at the long-wave surface)
and thus experience effective gravitational acceleration that varies in space
and time \citep{phillips1981dispersion,longuet1986eulerian,longuet1987propagation}.
Inserting (\ref{eq:dispersion}) into (\ref{eq:wave_crests}) and noting that
$\partial \sqrt{gk} / \partial x = 1/2 \sqrt{g/k} \partial k / \partial x + 1/2 \sqrt{k/g} \partial g / \partial x$,
we get:

@@

We now describe the analytical solutions for the short-wave wavenumber,
effective gravitational acceleration, amplitude, steepness, intrinsic frequency,
and intrinsic phase speed in presence of long waves.
The wavenumber is derived by linearising the conservation of wave crests
(\ref{eq:wave_crests}) and neglecting the spatial gradients of $k$.
The effective gravity is derived by projecting the long-wave induced surface
accelerations onto the local vertical axis (that is, the axis normal to the
long-wave surface), akin to \citet{zhang1990evolution}.
The amplitude is derived by first evaluating the short-wave action from the
action conservation (\ref{eq:wave_action}) and then applying the modulated
effective gravity to obtain the amplitude.
From the modulated wavenumber, amplitude, and effective gravity, the steepness,
frequency, and phase speed readily follow from the usual wave relationships.
Going forward we will use the tilde over the variables to denote the modulated
quantities (e.g. $\widetilde{k}$ for wavenumber).

@@

\backsection[Funding]{
  Milan Curcic was partly supported by the National Science Foundation
  grant AGS-1745384, Transatlantic Research Partnership, and the Office of Naval
  Research Coastal Land Air-Sea Interaction DRI.
}

>>>>

\begin{abstract}
Hydrodynamic modulation of short ocean surface waves by longer ambient waves is
a well-known ocean surface process that affects remote sensing, the
interpretation of \textit{in situ} wave measurements, and numerical wave
forecasting.
In this paper, we revisit the linear wave theory and derive higher-order
steady solutions for the change of short-wave wavenumber, action density, and
gravitational acceleration due to the presence of longer waves.
We validate the analytical solutions with numerical simulations of the full wave
crest and action conservation equations.
The nonlinear analytical solutions of short-wave wavenumber, amplitude, and
steepness modulation significantly deviate from the linear analytical solutions
of \citet{longuet1960changes}, and are similar to the nonlinear numerical
solutions by \citet{longuet1987propagation} and \citet{zhang1990evolution}.
The short-wave steepness modulation is attributed to be 5/8 due to the
wavenumber, 1/4 due to the wave action, and 1/8 due to the effective gravity.
We further examine the result of \citet{peureux2021unsteady} who found through
numerical simulations that the short-wave modulation grows unsteadily with
each long-wave passage.
We show that this unsteady growth only occurs only for homogeneous initial
conditions as a special case and does not generally occur, for example in more
realistic long-wave groups.
The proposed steady solutions are a good approximation of the fully
nonlinear numerical solutions in long-wave steepness up to $\approx$0.2.
Except for a subset of initial conditions, the solutions to the fully nonlinear
crest-action conservation equations are mostly steady in the reference frame of
the long waves.
\end{abstract}

@@

Short waves that ride on the surface of the longer waves move in an accelerated
reference frame (due to the centripetal acceleration at the long-wave surface)
and thus experience effective gravitational acceleration that varies in space
and time \citep{phillips1981dispersion,longuet1986eulerian,longuet1987propagation}.
Inserting (\ref{eq:dispersion}) into (\ref{eq:wave_crests}) and noting that:

\begin{equation}
\label{eq:dispersion_derivative}
\dfrac{\partial \sqrt{gk}}{\partial x} =
\dfrac{1}{2} \sqrt{\dfrac{g}{k}} \dfrac{\partial k}{\partial x} + \dfrac{1}{2} \sqrt{\dfrac{k}{g}} \dfrac{\partial g}{\partial x}
\end{equation}
we get:

@@

We now describe the analytical solutions for the short-wave wavenumber,
effective gravitational acceleration, amplitude, steepness, intrinsic frequency,
and intrinsic phase speed in presence of long waves.
The wavenumber is derived by linearising the conservation of wave crests
(\ref{eq:wave_crests}) and neglecting the spatial gradients of $k$ and $g$.
The effective gravity is derived by projecting the long-wave induced surface
accelerations onto the local vertical axis (that is, the axis normal to the
long-wave surface), akin to \citet{zhang1990evolution}.
The amplitude is derived by first evaluating the short-wave action from the
action conservation (\ref{eq:wave_action}) and then applying the modulated
effective gravity to obtain the amplitude.
From the modulated wavenumber, amplitude, and effective gravity, the steepness,
frequency, and phase speed readily follow from the usual wave relationships.
Going forward we will use the tilde over the variables to denote the modulated
quantities (\textit{e.g.} $\widetilde{k}$ for wavenumber).

@@

\backsection[Funding]{
  Milan Curcic was partly supported by the National Science Foundation
  grant AGS-1745384, Transatlantic Research Partnership by the French Embassy
  in the United States and the FACE Foundation, and the Office of Naval Research
  Coastal Land Air-Sea Interaction DRI.
}
